---
title: "Stress Physiology"
author: "jillashey"
date: "7/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(car)
library(rcompanion)
library(gdata)
library(multcomp)
library(nlme)
library(MASS)
library(lme4)
library(FSA)
library(tidyverse)
library(devtools)
library(gridExtra)
library(plotrix)
library(ggpubr)
```

```{r}
phys <- read.csv("data/FRAP/HT_FRAP_bothSets_042320.csv", header = T)
head(phys)
phys$Site <- as.factor(phys$Site)
phys$Treatment <- as.factor(phys$Treatment)
set1 <- filter(phys, Set  == "Set1")
set2 <- filter(phys, Set  == "Set2")
```

# Set1
# Evaluate normality for all variables 

Evaluate normality using histograms, Levene's test, Bartlett test, and QQplots.
- Levene's test: more robust to departures from normality than Bartlett's test; p>0.05 means the variances are equal
- Bartlett test: If the data is normally distributed, this is the best test to use. It is more likely to return a "false positive" when the data is non-normal; p>0.05 means the variances are equal
- Q-Q plot: tests the hypothesis that the data is normally distributed.  If the points on the plot are reasonably well approximated by a straight line, the popular Gaussian data hypothesis is plausible, while marked deviations from linearity provide evidence against this hypothesis and these data points should be removed. 

```{r}

```








