---
title: "Field WQ"
author: "jillashey"
date: "7/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(car)
library(rcompanion)
library(gdata)
library(multcomp)
library(nlme)
library(MASS)
library(lme4)
library(FSA)
library(tidyverse)
```

# Read in data 
```{r}
field_WQ <- read.csv("data/WQ/HT_FieldWaterQuality.csv", header = T)
head(field_WQ)
```

# Evaluate normality for all WQ variables 

Evaluate normality using histograms, Levene's test, Bartlett test, and QQplots.
- Levene's test: more robust to departures from normality than Bartlett's test; p>0.05 means the variances are equal
- Bartlett test: If the data is normally distributed, this is the best test to use. It is more likely to return a "false positive" when the data is non-normal; p>0.05 means the variances are equal
- Q-Q plot: tests the hypothesis that the data is normally distributed.  If the points on the plot are reasonably well approximated by a straight line, the popular Gaussian data hypothesis is plausible, while marked deviations from linearity provide evidence against this hypothesis and these data points should be removed.  

```{r}
## Temperature 
# Evaluate normality
hist(field_WQ$Temperature)
plotNormalHistogram(field_WQ$Temperature)
leveneTest(Temperature~Site, data=field_WQ) #normal
bartlett.test(Temperature~Site, data=field_WQ) #normal
fit<-lm(Temperature~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
t_log = log(field_WQ$Temperature)
plotNormalHistogram(t_log)
t_sqrt = sqrt(field_WQ$Temperature)
plotNormalHistogram(t_sqrt) 
t_cube= sign(field_WQ$Temperature) * abs(field_WQ$Temperature)^(1/3)
plotNormalHistogram(t_cube) 
t_recp = 1/(field_WQ$Temperature)
plotNormalHistogram(t_recp)
# Based on initial histograms, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
field_WQ = cbind(field_WQ, t_sqrt, t_cube)

# Evaluate normality in sqrt transformed temperature 
leveneTest(t_sqrt~Site, data=field_WQ) #normal
fit<-lm(t_sqrt~Site, data=field_WQ)
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
#similar s-shape to the untransformed variable

# Evaluate normality in cubed transformed temperature 
leveneTest(t_cube~Site, data=field_WQ) #normal
fit<-lm(t_cube~Site, data=field_WQ)
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## Not enough justification to transform - NO TRANSFORM in temperature
```

```{r}
## Salinity 
# Evaluate normality
hist(field_WQ$Salinity)
plotNormalHistogram(field_WQ$Salinity)
leveneTest(Salinity~Site, data=field_WQ) #normal
bartlett.test(Salinity~Site, data=field_WQ) #not normal
fit<-lm(Salinity~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
Slog = log(field_WQ$Salinity)
plotNormalHistogram(Slog)
Ssqrt = sqrt(field_WQ$Salinity)
plotNormalHistogram(Ssqrt)
Scube= sign(field_WQ$Salinity) * abs(field_WQ$Salinity)^(1/3)
plotNormalHistogram(Scube)
Srecp = 1/(field_WQ$Salinity)
plotNormalHistogram(Srecp)
# Based on initial histograms, log, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
field_WQ = cbind(field_WQ, Slog, Ssqrt, Scube)

# Evaluate normality in log transformed salinity 
leveneTest(Slog~Site, data=field_WQ) #normal - but barely
plotNormalHistogram(field_WQ$Slog)
bartlett.test(Slog~Site, data=field_WQ) #not normal
fit<-lm(Slog~Site, data=field_WQ)
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in sqrt transformed salinity 
leveneTest(Ssqrt~Site, data=field_WQ) #normal
plotNormalHistogram(field_WQ$Ssqrt)
bartlett.test(Ssqrt~Site, data=field_WQ) #not normal
fit<-lm(Ssqrt~Site, data=field_WQ)
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed salinity 
leveneTest(Scube~Site, data=field_WQ) #normal
plotNormalHistogram(field_WQ$Scube)
bartlett.test(Scube~Site, data=field_WQ) #not normal
fit<-lm(Scube~Site, data=field_WQ)
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## Not enough justification to transform - NO TRANSFORM in salinity
```

```{r}
## Spec pH
# Evaluate normality
hist(field_WQ$specpH)
plotNormalHistogram(field_WQ$specpH)
leveneTest(specpH~Site, data=field_WQ) # normal
bartlett.test(specpH~Site, data=field_WQ) # normal 
fit<-lm(specpH~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
##pretty good fit all things considered 

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
pHlog = log(field_WQ$specpH) 
plotNormalHistogram(pHlog)
pHsqrt = sqrt(field_WQ$specpH) 
plotNormalHistogram(pHsqrt)
pHcube= sign(field_WQ$specpH) * abs(field_WQ$specpH)^(1/3)
plotNormalHistogram(pHcube)
pHrecp = 1/(field_WQ$specpH) 
plotNormalHistogram(pHrecp)
# Based on initial histograms, log, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
field_WQ = cbind(field_WQ, pHlog, pHsqrt, pHcube)

# Evaluate normality in log transformed spec pH 
leveneTest(pHlog~Site, data=field_WQ) # normal
bartlett.test(pHlog~Site, data=field_WQ) # not normal 
fit<-lm(pHlog~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in sqrt transformed spec pH 
leveneTest(pHsqrt~Site, data=field_WQ) # normal
bartlett.test(pHsqrt~Site, data=field_WQ) # normal 
fit<-lm(pHsqrt~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed spec pH 
leveneTest(pHcube~Site, data=field_WQ) # normal
bartlett.test(pHcube~Site, data=field_WQ) # not normal - but barely 
fit<-lm(pHcube~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## Not enough justification to transform - NO TRANSFORM in spec pH
```

```{r}
## DO mg/L
hist(field_WQ$DO_mg)
plotNormalHistogram(field_WQ$DO_mg)
leveneTest(DO_mg~Site, data=field_WQ) # normal
bartlett.test(DO_mg~Site, data=field_WQ) # normal 
fit<-lm(DO_mg~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
DOlog = log(field_WQ$DO_mg) 
plotNormalHistogram(DOlog)
DOsqrt = sqrt(field_WQ$DO_mg) 
plotNormalHistogram(DOsqrt) 
DOcube= sign(field_WQ$DO_mg) * abs(field_WQ$DO_mg)^(1/3) 
plotNormalHistogram(DOcube) 
DOrecp = 1/(field_WQ$DO_mg) 
plotNormalHistogram(DOrecp)
# Based on initial histograms, log, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set
field_WQ = cbind(field_WQ, DOlog, DOsqrt, DOcube)

# Evaluate normality in log transformed spec pH 
leveneTest(DOlog~Site, data=field_WQ) # normal
bartlett.test(DOlog~Site, data=field_WQ) # normal 
fit<-lm(DOlog~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in sqrt transformed spec pH 
leveneTest(DOsqrt~Site, data=field_WQ) # normal
bartlett.test(DOsqrt~Site, data=field_WQ) # normal 
fit<-lm(DOsqrt~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed spec pH 
leveneTest(DOcube~Site, data=field_WQ) # normal
bartlett.test(DOcube~Site, data=field_WQ) # normal
fit<-lm(DOcube~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## Not enough justification to transform - NO TRANSFORM in DO_mg
```

```{r}
## Total alkalinity 
hist(field_WQ$TA)
plotNormalHistogram(field_WQ$TA)
leveneTest(TA~Site, data=field_WQ) # normal
bartlett.test(TA~Site, data=field_WQ) # not normal
fit<-lm(TA~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
#decent fit

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
TAlog = log(field_WQ$TA) 
plotNormalHistogram(TAlog)
TAsqrt = sqrt(field_WQ$TA) 
plotNormalHistogram(TAsqrt) 
TAcube = sign(field_WQ$TA) * abs(field_WQ$TA)^(1/3) 
plotNormalHistogram(TAcube) 
TArecp = 1/(field_WQ$TA) 
plotNormalHistogram(TArecp)
# Based on initial histograms, log, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set
field_WQ = cbind(field_WQ, TAlog, TAsqrt, TAcube)

# Evaluate normality in log transformed TA
leveneTest(TAlog~Site, data=field_WQ) # normal
bartlett.test(TAlog~Site, data=field_WQ) # not normal 
fit<-lm(TAlog~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in sqrt transformed TA 
leveneTest(TAsqrt~Site, data=field_WQ) # normal
bartlett.test(TAsqrt~Site, data=field_WQ) # not normal 
fit<-lm(TAsqrt~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed TA 
leveneTest(TAcube~Site, data=field_WQ) # normal
bartlett.test(TAcube~Site, data=field_WQ) # not normal
fit<-lm(TAcube~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## Not enough justification to transform - NO TRANSFORM in TA
```

```{r}
###pCO2
hist(field_WQ$pCO2)
plotNormalHistogram(field_WQ$pCO2)
leveneTest(pCO2~Site, data=field_WQ) # normal
bartlett.test(pCO2~Site, data=field_WQ) # not normal
fit<-lm(pCO2~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
pCO2log = log(field_WQ$pCO2) 
plotNormalHistogram(pCO2log) 
pCO2sqrt = sqrt(field_WQ$pCO2) 
plotNormalHistogram(pCO2sqrt) 
pCO2cube= sign(field_WQ$pCO2) * abs(field_WQ$pCO2)^(1/3) 
plotNormalHistogram(pCO2cube) 
pCO2recp = 1/(field_WQ$pCO2) 
plotNormalHistogram(pCO2recp) 
# Based on initial histograms, log, cube, and recp look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set
field_WQ = cbind(field_WQ, pCO2log, pCO2cube, pCO2recp)

# Evaluate normality in log transformed pCO2
leveneTest(pCO2log~Site, data=field_WQ) # normal
bartlett.test(pCO2log~Site, data=field_WQ) # normal 
fit<-lm(pCO2log~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed pCO2
leveneTest(pCO2cube~Site, data=field_WQ) # normal
bartlett.test(pCO2cube~Site, data=field_WQ) # normal 
fit<-lm(pCO2cube~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in recp transformed pCO2
leveneTest(pCO2recp~Site, data=field_WQ) # normal
bartlett.test(pCO2recp~Site, data=field_WQ) # normal 
fit<-lm(pCO2recp~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

## ????
```

```{r}
## Omega Ca
hist(field_WQ$Omega_Ca)
plotNormalHistogram(field_WQ$Omega_Ca)
leveneTest(Omega_Ca~Site, data=field_WQ) # normal
bartlett.test(Omega_Ar~Site, data=field_WQ) # not normal
fit<-lm(Omega_Ca~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
Calog = log(field_WQ$Omega_Ca) 
plotNormalHistogram(Calog) 
Casqrt = sqrt(field_WQ$Omega_Ca) 
plotNormalHistogram(Casqrt) 
Cacube= sign(field_WQ$Omega_Ca) * abs(field_WQ$Omega_Ca)^(1/3) 
plotNormalHistogram(Cacube) 
Carecp = 1/(field_WQ$Omega_Ca) 
plotNormalHistogram(Carecp) 
# Based on initial histograms, log, sqrt, and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set
field_WQ = cbind(field_WQ, Calog, Casqrt, Cacube)

# Evaluate normality in log transformed OmegaCa
leveneTest(Calog~Site, data=field_WQ) # normal
bartlett.test(Calog~Site, data=field_WQ) # not normal 
fit<-lm(Calog~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in log transformed OmegaCa
leveneTest(Calog~Site, data=field_WQ) # normal
bartlett.test(Calog~Site, data=field_WQ) # not normal 
fit<-lm(Calog~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in sqrt transformed OmegaCa
leveneTest(Casqrt~Site, data=field_WQ) # normal
bartlett.test(Casqrt~Site, data=field_WQ) # not normal 
fit<-lm(Casqrt~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed OmegaCa
leveneTest(Cacube~Site, data=field_WQ) # normal
bartlett.test(Cacube~Site, data=field_WQ) # not normal 
fit<-lm(Cacube~Site, data=field_WQ) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# ????
```

# Check if data contains any major outliers
Identify outliers with Bonferroni test
- Bonferroni test: null for the Bonferroni adjusted outlier test is the observation of an outlier. p<0.05 means no outliers.
```{r}
fit_temp<-lm(Temperature~Site, data=field_WQ)
outlierTest(fit_temp) # no outliers - unadj p > 0.05 & Bonnferroni p NA

fit_sal<-lm(Salinity~Site, data=field_WQ)
outlierTest(fit_sal) # potential outliers? unadj p < 0.05 & Bonnferroni p > 0.05

fit_spec<-lm(specpH~Site, data=field_WQ)
outlierTest(fit_spec) # potential outliers? unadj p < 0.05 & Bonnferroni p < 0.05 

fit_DOmg<-lm(DO_mg~Site, data=field_WQ)
outlierTest(fit_DOmg) # potential outliers? unadj p < 0.05 & Bonnferroni p > 0.05 

fit_TA<-lm(TA~Site, data=field_WQ)
outlierTest(fit_TA) # potential outliers? unadj p < 0.05 & Bonnferroni p < 0.05

fit_pCO2<-lm(pCO2~Site, data=field_WQ)
outlierTest(fit_pCO2) # potential outliers? unadj p < 0.05 & Bonnferroni p < 0.05 

fit_Ca<-lm(Omega_Ca~Site, data=field_WQ)
outlierTest(fit_Ca) # potential outliers? unadj p < 0.05 & Bonnferroni p < 0.05 

```

# Run one-way ANOVAs to identify differences by site 
#### Comparing distal sites: MR, PR, UC, VP
```{r}
#field_WQ$Site<-as.factor(field_WQ$Site)

# Filter by site 
MR <- filter(field_WQ, Site  == "MR")
PR <- filter(field_WQ, Site  == "PR")
UC <- filter(field_WQ, Site  == "UC")
VP <- filter(field_WQ, Site  == "VP")
VP_2 <- filter(field_WQ, Site  == "VP-2")

# Bind distal sites together in df 
AllSite <- rbind(MR,PR,UC,VP)
AllSite$Site <- as.factor(AllSite$Site)
```

## Compare field water quality by distal sites
```{r}
## Temperature 

# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Temperature~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # none significant
Anova(fit2, type="III") # not significant
kruskal.test(Temperature~Site, data=AllSite) # not significant

# Second, I will do the ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed (sqrt and cube)
# Sqrt
fit2 <-lm(t_sqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # none significant
Anova(fit2, type="III") # not significant
kruskal.test(t_sqrt~Site, data=AllSite) # not significant

# Cube
fit2 <-lm(t_cube~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # none significant
Anova(fit2, type="III") # not significant
kruskal.test(t_cube~Site, data=AllSite) # not significant
```

```{r}
## Salinity 
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Salinity~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # all comparisons significant
Anova(fit2, type="III") # significant
test_sal <- glht(fit2, linfct = mcp(Site = "Tukey")) # post-hoc test
summary(test_sal) # significant comparisons by site 
kruskal.test(Salinity~Site, data=AllSite) # signicant 
dunnTest(Salinity~Site, data=AllSite) # Automatically fills in Holm method; significant comparisons by site

# Second, I will do the ANOVA test using the transformed variables (log, sqrt, cube) to compare/check them to the untransformed 
# Log
fit2 <-lm(Slog~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # all comparisons significant
Anova(fit2, type="III") # significant
test_sal <- glht(fit2, linfct = mcp(Site = "Tukey")) # post-hoc test
summary(test_sal) # significant comparisons by site
kruskal.test(Slog~Site, data=AllSite) # significant
dunnTest(Slog~Site, data=AllSite) # Automatically fills in Holm method; significant comparisons by site

# Sqrt
fit2 <-lm(Ssqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # all comparisons significant
Anova(fit2, type="III") # significant
test_sal <- glht(fit2, linfct = mcp(Site = "Tukey")) # post-hoc test
summary(test_sal) # significant comparisons by site
kruskal.test(Ssqrt~Site, data=AllSite) # significant
dunnTest(Ssqrt~Site, data=AllSite) # Automatically fills in Holm method; significant comparisons by site

# Cube
fit2 <-lm(Scube~Site, na.action = na.exclude, data=AllSite)
summary(fit2) # all comparisons significant
Anova(fit2, type="III") # significant
test_sal <- glht(fit2, linfct = mcp(Site = "Tukey")) # post-hoc test
summary(test_sal) # significant comparisons by site
kruskal.test(Scube~Site, data=AllSite) # significant
dunnTest(Scube~Site, data=AllSite) # Automatically fills in Holm method; significant comparisons by site
```

```{r}
## spec pH
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(specpH~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(specpH~Site, data=AllSite) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pHlog~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(pHlog~Site, data=AllSite) # not significant

# Sqrt
fit2 <-lm(pHsqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pHsqrt~Site, data=AllSite) # not significant

# Cube
fit2 <-lm(pHcube~Site, na.action = na.exclude, data=AllSite) 
summary(fit2) # PR signficant
Anova(fit2, type="III") # not significant 
kruskal.test(pHcube~Site, data=AllSite) # not significant
```

```{r}
## DO mg
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(DO_mg~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(DO_mg~Site, data=AllSite) # not significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(DOlog~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(DOlog~Site, data=AllSite) # not significant

# Log
fit2 <-lm(DOsqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(DOsqrt~Site, data=AllSite) # not significant

# Cube
fit2 <-lm(DOcube~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(DOcube~Site, data=AllSite) # not significant
```

```{r}
## TA
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(TA~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant
test <- glht(fit2, linfct = mcp(Site = "Tukey")) 
summary(test) # significant comparisons by site
kruskal.test(TA~Site, data=AllSite) # significant
dunnTest(TA~Site, data=AllSite) # significant comparisons by site

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(TAlog~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significiant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # significant comparisons by site
kruskal.test(TAlog~Site, data=AllSite) # significant
dunnTest(TAlog~Site, data=AllSite) # significant comparisons by site

# Sqrt
fit2 <-lm(TAsqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # significant comparisons by site
kruskal.test(TAsqrt~Site, data=AllSite) # significant
dunnTest(TAsqrt~Site, data=AllSite) # significant comparisons by site

# Cube
fit2 <-lm(TAcube~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant 
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # significant comparisons by site
kruskal.test(TAcube~Site, data=AllSite) # significant
dunnTest(TAcube~Site, data=AllSite) # significant comparisons by site
```

```{r}
# pCO2
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(pCO2~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # no significant comparisons by site 
kruskal.test(pCO2~Site, data=AllSite) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pCO2log~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2log~Site, data=AllSite) # not significant

# Cube
fit2 <-lm(pCO2cube~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # no significant comparisons by site 
kruskal.test(pCO2cube~Site, data=AllSite) # not significant

# Reciprocal
fit2 <-lm(pCO2recp~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(pCO2recp~Site, data=AllSite) # not significant
```

```{r}
##Omega Ca
##First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Omega_Ca~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Omega_Ca~Site, data=AllSite) # significant
dunnTest(Omega_Ca~Site, data=AllSite) # significant comparisons by site

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(Calog~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # significant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # significant comparisons by site
kruskal.test(Calog~Site, data=AllSite) # significant
dunnTest(Calog~Site, data=AllSite) # significant comparisons by site

# Sqrt
fit2 <-lm(Casqrt~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Casqrt~Site, data=AllSite) # significant
dunnTest(Casqrt~Site, data=AllSite) # significant comparisons by site

# Cube
fit2 <-lm(Cacube~Site, na.action = na.exclude, data=AllSite)
summary(fit2)
Anova(fit2, type="III") # slightly significant
test <- glht(fit2, linfct = mcp(Site = "Tukey"))
summary(test) # significant comparisons by site
kruskal.test(Cacube~Site, data=AllSite) # significant
dunnTest(Cacube~Site, data=AllSite) # significant comparisons by site
```


#### Comparing MR and VP-2
```{r}
# Bind sites together in df 
MR_VP2 <- rbind(MR,VP_2)
MR_VP2$Site <- as.factor(MR_VP2$Site)
```

## Compare field water quality by MR and VP-2 sites

```{r}
## Temperature 

# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Temperature~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Temperature~Site, data=MR_VP2) # significant

# Second, I will do the ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed (sqrt and cube)
# Sqrt
fit2 <-lm(t_sqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_sqrt~Site, data=MR_VP2) # significant

# Cube
fit2 <-lm(t_cube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_cube~Site, data=MR_VP2) # significant
```

```{r}
## Salinity 
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Salinity~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Salinity~Site, data=MR_VP2) # significant 

# Second, I will do the ANOVA test using the transformed variables (log, sqrt, cube) to compare/check them to the untransformed 
# Log
fit2 <-lm(Slog~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Slog~Site, data=MR_VP2) # significant

# Sqrt
fit2 <-lm(Ssqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Ssqrt~Site, data=MR_VP2) # significant

# Cube
fit2 <-lm(Scube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Scube~Site, data=MR_VP2) # significant
```

```{r}
## spec pH
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(specpH~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(specpH~Site, data=MR_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pHlog~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant 
kruskal.test(pHlog~Site, data=MR_VP2) # not significant

# Sqrt
fit2 <-lm(pHsqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pHsqrt~Site, data=MR_VP2) # not significant

# Cube
fit2 <-lm(pHcube~Site, na.action = na.exclude, data=MR_VP2) 
summary(fit2) 
Anova(fit2, type="III") # significant 
kruskal.test(pHcube~Site, data=MR_VP2) # not significant
```

```{r}
## DO mg
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(DO_mg~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DO_mg~Site, data=MR_VP2) # significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(DOlog~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOlog~Site, data=MR_VP2) # significant

# Log
fit2 <-lm(DOsqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOsqrt~Site, data=MR_VP2) # significant

# Cube
fit2 <-lm(DOcube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOcube~Site, data=MR_VP2) # significant
```

```{r}
## TA
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(TA~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TA~Site, data=MR_VP2) # not significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(TAlog~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significiant
kruskal.test(TAlog~Site, data=MR_VP2) # not significant

# Sqrt
fit2 <-lm(TAsqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TAsqrt~Site, data=MR_VP2) # not significant

# Cube
fit2 <-lm(TAcube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(TAcube~Site, data=MR_VP2) # not significant
```

```{r}
# pCO2
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(pCO2~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2~Site, data=MR_VP2) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pCO2log~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2log~Site, data=MR_VP2) # significant

# Cube
fit2 <-lm(pCO2cube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2cube~Site, data=MR_VP2) # significant

# Reciprocal
fit2 <-lm(pCO2recp~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant 
kruskal.test(pCO2recp~Site, data=MR_VP2) # significant
```

```{r}
##Omega Ca
##First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Omega_Ca~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Omega_Ca~Site, data=AllSite) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(Calog~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Calog~Site, data=MR_VP2) # not significant

# Sqrt
fit2 <-lm(Casqrt~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Casqrt~Site, data=MR_VP2) # significant

# Cube
fit2 <-lm(Cacube~Site, na.action = na.exclude, data=MR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Cacube~Site, data=MR_VP2) # not significant
```


#### Comparing PR and VP-2
```{r}
# Bind sites together in df 
PR_VP2 <- rbind(PR,VP_2)
PR_VP2$Site <- as.factor(PR_VP2$Site)
```

## Compare field water quality by MR and VP-2 sites

```{r}
## Temperature 

# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Temperature~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Temperature~Site, data=PR_VP2) # significant

# Second, I will do the ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed (sqrt and cube)
# Sqrt
fit2 <-lm(t_sqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_sqrt~Site, data=PR_VP2) # significant

# Cube
fit2 <-lm(t_cube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_cube~Site, data=PR_VP2) # significant
```

```{r}
## Salinity 
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Salinity~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Salinity~Site, data=PR_VP2) # significant 

# Second, I will do the ANOVA test using the transformed variables (log, sqrt, cube) to compare/check them to the untransformed 
# Log
fit2 <-lm(Slog~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Slog~Site, data=PR_VP2) # significant

# Sqrt
fit2 <-lm(Ssqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Ssqrt~Site, data=PR_VP2) # significant

# Cube
fit2 <-lm(Scube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Scube~Site, data=PR_VP2) # significant
```

```{r}
## spec pH
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(specpH~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(specpH~Site, data=PR_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pHlog~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant 
kruskal.test(pHlog~Site, data=PR_VP2) # not significant

# Sqrt
fit2 <-lm(pHsqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pHsqrt~Site, data=PR_VP2) # not significant

# Cube
fit2 <-lm(pHcube~Site, na.action = na.exclude, data=PR_VP2) 
summary(fit2) 
Anova(fit2, type="III") # significant 
kruskal.test(pHcube~Site, data=PR_VP2) # not significant
```

```{r}
## DO mg
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(DO_mg~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DO_mg~Site, data=PR_VP2) # significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(DOlog~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOlog~Site, data=PR_VP2) # significant

# Log
fit2 <-lm(DOsqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOsqrt~Site, data=PR_VP2) # significant

# Cube
fit2 <-lm(DOcube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOcube~Site, data=PR_VP2) # significant
```

```{r}
## TA
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(TA~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TA~Site, data=PR_VP2) # not significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(TAlog~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significiant
kruskal.test(TAlog~Site, data=PR_VP2) # not significant

# Sqrt
fit2 <-lm(TAsqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TAsqrt~Site, data=PR_VP2) # not significant

# Cube
fit2 <-lm(TAcube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(TAcube~Site, data=PR_VP2) # not significant
```

```{r}
# pCO2
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(pCO2~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2~Site, data=PR_VP2) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pCO2log~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2log~Site, data=PR_VP2) # significant

# Cube
fit2 <-lm(pCO2cube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(pCO2cube~Site, data=PR_VP2) # significant

# Reciprocal
fit2 <-lm(pCO2recp~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant 
kruskal.test(pCO2recp~Site, data=PR_VP2) # significant
```

```{r}
##Omega Ca
##First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Omega_Ca~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Omega_Ca~Site, data=PR_VP2) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(Calog~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Calog~Site, data=PR_VP2) # not significant

# Sqrt
fit2 <-lm(Casqrt~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Casqrt~Site, data=PR_VP2) # significant

# Cube
fit2 <-lm(Cacube~Site, na.action = na.exclude, data=PR_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Cacube~Site, data=PR_VP2) # not significant
```



#### Comparing UC and VP-2
```{r}
# Bind sites together in df 
UC_VP2 <- rbind(UC,VP_2)
UC_VP2$Site <- as.factor(UC_VP2$Site)
```

## Compare field water quality by UC and VP-2 sites

```{r}
## Temperature 

# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Temperature~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Temperature~Site, data=UC_VP2) # significant

# Second, I will do the ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed (sqrt and cube)
# Sqrt
fit2 <-lm(t_sqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_sqrt~Site, data=UC_VP2) # significant

# Cube
fit2 <-lm(t_cube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_cube~Site, data=UC_VP2) # significant
```

```{r}
## Salinity 
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Salinity~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Salinity~Site, data=UC_VP2) # significant 

# Second, I will do the ANOVA test using the transformed variables (log, sqrt, cube) to compare/check them to the untransformed 
# Log
fit2 <-lm(Slog~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Slog~Site, data=UC_VP2) # significant

# Sqrt
fit2 <-lm(Ssqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Ssqrt~Site, data=UC_VP2) # significant

# Cube
fit2 <-lm(Scube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Scube~Site, data=UC_VP2) # significant
```

```{r}
## spec pH
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(specpH~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(specpH~Site, data=UC_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pHlog~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(pHlog~Site, data=UC_VP2) # not significant

# Sqrt
fit2 <-lm(pHsqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pHsqrt~Site, data=UC_VP2) # not significant

# Cube
fit2 <-lm(pHcube~Site, na.action = na.exclude, data=UC_VP2) 
summary(fit2) 
Anova(fit2, type="III") # not significant 
kruskal.test(pHcube~Site, data=UC_VP2) # not significant
```

```{r}
## DO mg
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(DO_mg~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DO_mg~Site, data=UC_VP2) # significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(DOlog~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOlog~Site, data=UC_VP2) # significant

# Log
fit2 <-lm(DOsqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOsqrt~Site, data=UC_VP2) # significant

# Cube
fit2 <-lm(DOcube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOcube~Site, data=UC_VP2) # significant
```

```{r}
## TA
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(TA~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TA~Site, data=UC_VP2) # not significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(TAlog~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significiant
kruskal.test(TAlog~Site, data=UC_VP2) # not significant

# Sqrt
fit2 <-lm(TAsqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TAsqrt~Site, data=UC_VP2) # not significant

# Cube
fit2 <-lm(TAcube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(TAcube~Site, data=UC_VP2) # not significant
```

```{r}
# pCO2
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(pCO2~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2~Site, data=UC_VP2) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pCO2log~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2log~Site, data=UC_VP2) # significant

# Cube
fit2 <-lm(pCO2cube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2cube~Site, data=UC_VP2) # significant

# Reciprocal
fit2 <-lm(pCO2recp~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # slightly significant 
kruskal.test(pCO2recp~Site, data=UC_VP2) # significant
```

```{r}
##Omega Ca
##First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Omega_Ca~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Omega_Ca~Site, data=UC_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(Calog~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Calog~Site, data=UC_VP2) # not significant

# Sqrt
fit2 <-lm(Casqrt~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Casqrt~Site, data=UC_VP2) # not significant

# Cube
fit2 <-lm(Cacube~Site, na.action = na.exclude, data=UC_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Cacube~Site, data=UC_VP2) # not significant
```


#### Comparing VP and VP-2
```{r}
# Bind sites together in df 
VP_VP2 <- rbind(VP,VP_2)
VP_VP2$Site <- as.factor(VP_VP2$Site)
```

## Compare field water quality by VP and VP-2 sites
```{r}
## Temperature 

# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Temperature~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Temperature~Site, data=VP_VP2) # significant

# Second, I will do the ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed (sqrt and cube)
# Sqrt
fit2 <-lm(t_sqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_sqrt~Site, data=VP_VP2) # significant

# Cube
fit2 <-lm(t_cube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(t_cube~Site, data=VP_VP2) # significant
```

```{r}
## Salinity 
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Salinity~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Salinity~Site, data=VP_VP2) # significant 

# Second, I will do the ANOVA test using the transformed variables (log, sqrt, cube) to compare/check them to the untransformed 
# Log
fit2 <-lm(Slog~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Slog~Site, data=VP_VP2) # significant

# Sqrt
fit2 <-lm(Ssqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2) 
Anova(fit2, type="III") # significant
kruskal.test(Ssqrt~Site, data=VP_VP2) # significant

# Cube
fit2 <-lm(Scube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(Scube~Site, data=VP_VP2) # significant
```

```{r}
## spec pH
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(specpH~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(specpH~Site, data=VP_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pHlog~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(pHlog~Site, data=VP_VP2) # not significant

# Sqrt
fit2 <-lm(pHsqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pHsqrt~Site, data=VP_VP2) # not significant

# Cube
fit2 <-lm(pHcube~Site, na.action = na.exclude, data=VP_VP2) 
summary(fit2) 
Anova(fit2, type="III") # not significant 
kruskal.test(pHcube~Site, data=VP_VP2) # not significant
```

```{r}
## DO mg
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(DO_mg~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DO_mg~Site, data=VP_VP2) # significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(DOlog~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOlog~Site, data=VP_VP2) # significant

# Log
fit2 <-lm(DOsqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOsqrt~Site, data=VP_VP2) # significant

# Cube
fit2 <-lm(DOcube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # significant
kruskal.test(DOcube~Site, data=VP_VP2) # significant
```

```{r}
## TA
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(TA~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TA~Site, data=VP_VP2) # not significant

# Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(TAlog~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significiant
kruskal.test(TAlog~Site, data=VP_VP2) # not significant

# Sqrt
fit2 <-lm(TAsqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(TAsqrt~Site, data=VP_VP2) # not significant

# Cube
fit2 <-lm(TAcube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant 
kruskal.test(TAcube~Site, data=VP_VP2) # not significant
```

```{r}
# pCO2
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(pCO2~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2~Site, data=VP_VP2) # significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(pCO2log~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2log~Site, data=VP_VP2) # significant

# Cube
fit2 <-lm(pCO2cube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(pCO2cube~Site, data=VP_VP2) # significant

# Reciprocal
fit2 <-lm(pCO2recp~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # slightly significant 
kruskal.test(pCO2recp~Site, data=VP_VP2) # significant
```

```{r}
##Omega Ca
##First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Omega_Ca~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Omega_Ca~Site, data=VP_VP2) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Log
fit2 <-lm(Calog~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Calog~Site, data=VP_VP2) # not significant

# Sqrt
fit2 <-lm(Casqrt~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Casqrt~Site, data=VP_VP2) # not significant

# Cube
fit2 <-lm(Cacube~Site, na.action = na.exclude, data=VP_VP2)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Cacube~Site, data=VP_VP2) # not significant
```




# Plot water quality variables 
```{r}
# Format data 
field_WQ$Date <- parse_date(field_WQ$Date, "%m/%d/%y")
```

```{r}
## Plot WQ variables 
Temps <- ggplot(field_WQ, aes(x=Date, y=Temperature)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("Temperature (C)") +
  theme_linedraw()
Temps

Sal <- ggplot(field_WQ, aes(x=Date, y=Salinity)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("Salinity (psu)") +
  theme_linedraw()
Sal

pH <- ggplot(field_WQ, aes(x=Date, y=specpH)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("pH Total") +
  theme_linedraw()
pH

DO <- ggplot(field_WQ, aes(x=Date, y=DO_mg)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("Dissolved oxygen (mg/mL)") + 
  theme_linedraw()
DO

TA <- ggplot(field_WQ, aes(x=Date, y=TA)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("Total alkalinity (mol/kg)") +
  theme_linedraw()
TA

pCO2 <- ggplot(field_WQ, aes(x=Date, y=pCO2)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab("pCO2 (atm)") +
  theme_linedraw()
pCO2

Ca <- ggplot(field_WQ, aes(x=Date, y=Omega_Ca)) +
  geom_line(aes(color = Site), size = 1)+
  scale_color_manual(values = c("mediumpurple3", "darkorange2", "palegreen3", "dodgerblue2", "goldenrod2")) +
  ylab(" Calcite") +
  theme_linedraw()
Ca

```

```{r}
# Save plots
# as pdfs
ggsave("output/SuppFigures/FigureS2.pdf", plot = Temps, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS3.pdf", plot = Sal, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS4.pdf", plot = pH, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS5.pdf", plot = DO, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS6.pdf", plot = TA, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS7.pdf", plot = pCO2, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS8.pdf", plot = Ca, width = 8, height = 4)

# as pngs
ggsave("output/SuppFigures/FigureS2.png", plot = Temps, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS3.png", plot = Sal, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS4.png", plot = pH, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS5.png", plot = DO, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS6.png", plot = TA, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS7.png", plot = pCO2, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS8.png", plot = Ca, width = 8, height = 4)
```


