---
title: "Shell Height"
author: "jillashey"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(devtools)
library(ggplot2)
library(ggpubr)
library(gridExtra)
library(plotrix)
```

# Read in data 
```{r}
SH <- read.csv("data/Growth/HT_InitialShellHeight_022219.csv", header = T)

# Subset by distal shell heights 
distal.SH <- subset(SH, EnvHist == "Distal")
```

# Evaluate normality for initial shell height 

Evaluate normality using histograms, Levene's test, Bartlett test, and QQplots.
- Levene's test: more robust to departures from normality than Bartlett's test; p>0.05 means the variances are equal
- Bartlett test: If the data is normally distributed, this is the best test to use. It is more likely to return a "false positive" when the data is non-normal; p>0.05 means the variances are equal
- Q-Q plot: tests the hypothesis that the data is normally distributed.  If the points on the plot are reasonably well approximated by a straight line, the popular Gaussian data hypothesis is plausible, while marked deviations from linearity provide evidence against this hypothesis and these data points should be removed.  

```{r}
## Initial Shell Height
# Evaluate normality
hist(distal.SH$Initial_Shell)
plotNormalHistogram(distal.SH$Initial_Shell)
leveneTest(Initial_Shell~Site, data=distal.SH) # not normal
bartlett.test(Initial_Shell~Site, data=distal.SH) # not normal 
fit<-lm(Initial_Shell~Site, data=distal.SH) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
SH_log = log(distal.SH$Initial_Shell)
plotNormalHistogram(SH_log)
SH_sqrt = sqrt(distal.SH$Initial_Shell)
plotNormalHistogram(SH_sqrt) 
SH_cube= sign(distal.SH$Initial_Shell) * abs(distal.SH$Initial_Shell)^(1/3)
plotNormalHistogram(SH_cube) 
SH_recp = 1/(distal.SH$Initial_Shell)
plotNormalHistogram(SH_recp)
# Based on initial histograms, sqrt and cube look sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
distal.SH = cbind(distal.SH, SH_sqrt, SH_cube)

# Evaluate normality in sqrt transformed temperature 
leveneTest(SH_sqrt~Site, data=distal.SH) # not normal
bartlett.test(SH_sqrt~Site, data=distal.SH) # not normal 
fit<-lm(SH_sqrt~Site, data=distal.SH) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Evaluate normality in cube transformed temperature 
leveneTest(SH_cube~Site, data=distal.SH) # not normal
bartlett.test(SH_cube~Site, data=distal.SH) # not normal 
fit<-lm(SH_cube~Site, data=distal.SH) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
title("QQ-plot of residuals")
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
title("QQ-plot of standardized residuals") 
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
```

# Run one-way ANOVA to identify differences by site 
#### Comparing distal sites: MR, PR, UC, VP
```{r}
## Initial shell height
# First, I will do the ANOVA test and the Kruskal-Wallis test (both using the not transformed variable) to compare results 
fit2 <-lm(Initial_Shell~Site, na.action = na.exclude, data=distal.SH)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(Initial_Shell~Site, data=distal.SH) # not significant

##Second, I will do the ANOVA test using the transformed variables to compare them to the untransformed 
# Sqrt
fit2 <-lm(SH_sqrt~Site, na.action = na.exclude, data=distal.SH)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(SH_sqrt~Site, data=distal.SH) # not significant

# Cube
fit2 <-lm(SH_cube~Site, na.action = na.exclude, data=distal.SH)
summary(fit2)
Anova(fit2, type="III") # not significant
kruskal.test(SH_cube~Site, data=distal.SH) # not significant
```

# Plot initial shell height 
```{r}
# Calculate mean and std error 
MR <- subset(distal.SH, Site=="Merroir")
MR_SH_mean <- mean(MR$Initial_Shell)
MR_SH_se <- std.error(MR$Initial_Shell)

PR <- subset(distal.SH, Site=="Piankatank")
PR_SH_mean = mean(PR$Initial_Shell)
PR_SH_se = std.error(PR$Initial_Shell)

UC <- subset(distal.SH, Site=="Urbanna Creek")
UC_SH_mean <- mean(UC$Initial_Shell)
UC_SH_se <- std.error(UC$Initial_Shell)

VP <- subset(distal.SH, Site=="VIMS Pier")
VP_SH_mean <- mean(VP$Initial_Shell)
VP_SH_se <- std.error(VP$Initial_Shell)

#make a data frame with the name (categorical variable - Site), the values (average of physiological measurements from each site),
#and se (standard error for each site)
distal.SH_summary = data.frame(name=c("Merroir", "Piankatank", "Urbanna Creek", "VIMS Pier"), 
                    value = c(MR_SH_mean, PR_SH_mean, UC_SH_mean, VP_SH_mean), # mean
                    se = c(MR_SH_se, PR_SH_se, UC_SH_se, VP_SH_se)) #std error

# Make bar plot for initial shell height 
p1 = ggplot(distal.SH_summary, aes(x=name, y=value)) + #establishing what's going on the plot
  geom_bar(stat = "identity", fill="gray", color = "black", size = 0.3, width = 0.5) + #its going to be a bar plot in these sizes, shapes, colors, etc
  geom_errorbar(aes(x=name, ymin = value - se, ymax = value + se), colour = "black", width = 0.1, alpha = 0.9, size = 0.5) + #error bars in specific size, shape, color, etc
  ylab("Initial Shell Height (mm)") + xlab("") + #x and y axes titles 
  #ggtitle("") + #title 
  scale_y_continuous(expand = c(0,0)) + #making the bars sit right on the x axis-- no white space bt the bars and the actual axis
  theme_minimal() + #setting the background theme
  theme(axis.line = element_line(), axis.text.x = element_text(size = 12, angle=30, hjust = 1, colour = "black"), axis.text.y = element_text(size=12, color = "black"), axis.title.y = element_text(size = 15)) #adding specific lines, sizes, degrees, etc for the axes texts and labels
p1

# Save plots 
ggsave("output/SuppFigures/FigureS1.pdf", plot = p1, width = 8, height = 4)
ggsave("output/SuppFigures/FigureS1.png", plot = p1, width = 8, height = 4)
```



