---
title: "Stress Physiology"
author: "jillashey"
date: "7/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(car)
library(rcompanion)
library(gdata)
library(multcomp)
library(nlme)
library(MASS)
library(lme4)
library(FSA)
library(tidyverse)
library(devtools)
library(gridExtra)
library(plotrix)
library(ggpubr)
```

```{r}
phys <- read.csv("data/FRAP/HT_FRAP_bothSets_042320.csv", header = T)
head(phys)
phys$Site <- as.factor(phys$Site)
phys$Treatment <- as.factor(phys$Treatment)
frap1 <- filter(phys, Set  == "Set1")
frap2 <- filter(phys, Set  == "Set2")
```

We will be using the test_FRAP variable (FRAP/g tissue sample) for analysis
# Set1
# Evaluate normality for all variables 

Evaluate normality using histograms, Levene's test, Bartlett test, and QQplots.
- Levene's test: more robust to departures from normality than Bartlett's test; p>0.05 means the variances are equal
- Bartlett test: If the data is normally distributed, this is the best test to use. It is more likely to return a "false positive" when the data is non-normal; p>0.05 means the variances are equal
- Q-Q plot: tests the hypothesis that the data is normally distributed.  If the points on the plot are reasonably well approximated by a straight line, the popular Gaussian data hypothesis is plausible, while marked deviations from linearity provide evidence against this hypothesis and these data points should be removed. 

```{r}
## FRAP
# Evaluate normality 
hist(frap1$test_FRAP)
plotNormalHistogram(frap1$test_FRAP)
leveneTest(test_FRAP ~interaction(Treatment, Site), data=frap1) #normal
bartlett.test(test_FRAP ~interaction(Treatment, Site), data=frap1) # normal
fit<-lm(test_FRAP ~interaction(Treatment, Site), data=frap1) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
# FRAP log and sqrt were previously calculated in df
plotNormalHistogram(frap1$Log_test_FRAP)
plotNormalHistogram(frap1$Sqrt_test_FRAP)
frap_cube1 = sign(frap1$test_FRAP) * abs(frap1$test_FRAP)^(1/3)
plotNormalHistogram(frap_cube1) 
frap_recp1 = 1/(frap1$test_FRAP)
plotNormalHistogram(frap_recp1)
# Based on initial histograms, recp looks sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
frap1 = cbind(frap1, frap_cube1)

# Log
leveneTest(Log_test_FRAP ~interaction(Treatment, Site), data=frap1) #normal
bartlett.test(Log_test_FRAP ~interaction(Treatment, Site), data=frap1) # not normal
fit<-lm(Log_test_FRAP ~interaction(Treatment, Site), data=frap1) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Sqrt
leveneTest(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap1) #normal
bartlett.test(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap1) # not normal
fit<-lm(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap1) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Cube
leveneTest(frap_cube1 ~interaction(Treatment, Site), data=frap1) #normal
bartlett.test(frap_cube1 ~interaction(Treatment, Site), data=frap1) # not normal
fit<-lm(frap_cube1 ~interaction(Treatment, Site), data=frap1) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
```

# Set 2
```{r}
## FRAP
# Evaluate normality 
hist(frap2$test_FRAP)
plotNormalHistogram(frap2$test_FRAP)
leveneTest(test_FRAP ~interaction(Treatment, Site), data=frap2) #normal
bartlett.test(test_FRAP ~interaction(Treatment, Site), data=frap2) # normal
fit<-lm(test_FRAP ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
# FRAP log and sqrt were previously calculated in df
plotNormalHistogram(frap2$Log_test_FRAP)
plotNormalHistogram(frap2$Sqrt_test_FRAP)
frap_cube2 = sign(frap2$test_FRAP) * abs(frap2$test_FRAP)^(1/3)
plotNormalHistogram(frap_cube2) 
frap_recp2 = 1/(frap2$test_FRAP)
plotNormalHistogram(frap_recp2)
# Based on initial histograms, sqrt and cube looks sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
frap2 = cbind(frap2, frap_cube2)

# Sqrt
leveneTest(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap2) #normal
bartlett.test(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap2) # not normal
fit<-lm(Sqrt_test_FRAP ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Cube
leveneTest(frap_cube2 ~interaction(Treatment, Site), data=frap2) #normal
bartlett.test(frap_cube2 ~interaction(Treatment, Site), data=frap2) # not normal
fit<-lm(frap_cube2 ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
```

```{r}
## Delta FRAP
# Evaluate normality 
hist(frap2$test_FRAP_Delta)
plotNormalHistogram(frap2$test_FRAP_Delta)
leveneTest(test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) #normal
bartlett.test(test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # normal
fit<-lm(test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Transform the variable using log, sqrt, cube, and reciprocal & look at histograms of transformed values
# FRAP log and sqrt were previously calculated in df
plotNormalHistogram(frap2$Log_test_FRAP_Delta)
plotNormalHistogram(frap2$Sqrt_test_FRAP_Delta)
d_frap_cube2 = sign(frap2$test_FRAP_Delta) * abs(frap2$test_FRAP_Delta)^(1/3)
plotNormalHistogram(d_frap_cube2) 
d_frap_recp2 = 1/(frap2$test_FRAP_Delta)
plotNormalHistogram(d_frap_recp2)
# Based on initial histograms, log and sqrt looks sort of normal. Will evaluate normality in these transformed values 

# Bind the transformed variables into the data set 
# none to bind 

# Log
leveneTest(Log_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # not normal
bartlett.test(Log_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # not normal
fit<-lm(Log_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")

# Sqrt
leveneTest(Sqrt_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) #normal
bartlett.test(Sqrt_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # not normal
fit<-lm(Sqrt_test_FRAP_Delta ~interaction(Treatment, Site), data=frap2) # qqplot
par(mfrow=c(2,2))
qqnorm(residuals(fit), ylab="Residuals")
qqline(residuals(fit))
qqnorm(rstandard(fit), ylab="Residuals")
qqline(rstandard(fit))
plot(fitted(fit), residuals(fit), xlab="Fitted", ylab="Residuals")
abline(h=0)
title("Residuals vs fitted")
plot(fitted(fit), abs(residuals(fit)), xlab="Fitted", ylab="Absolute residuals")
abline(h=0)
title("Absolute residuals vs fitted")
```

# Run two-way ANOVAs to identify differences by site 
#### Comparing sites (MR, PR, UC, VP) and Treatments (MR-A, PR-UC-A, VP-A, Low) for Set1 

```{r}
## FRAP - set 1

# First, I will do two-way ANOVA test ( using the untransformed variable)
# Make full model with treatment and site 
LME_FRAP <-lme(test_FRAP~Treatment+Site,random = ~1|Replicate, data=frap1)
Anova(LME_FRAP, type="III") ##site significant
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(test_FRAP~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)

# Second, I will do two-way ANOVA test using the transformed (log, sqrt and cube) variables to compare/check them to the untransformed
# Log 
# Make full model with treatment and site 
LME_FRAP <-lme(Log_test_FRAP~Treatment+Site,random = ~1|Replicate, data=frap1)
Anova(LME_FRAP, type="III") ##site significant
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(Log_test_FRAP~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)

# Sqrt 
# Make full model with treatment and site 
LME_FRAP <-lme(Sqrt_test_FRAP~Treatment+Site,random = ~1|Replicate, data=frap1)
Anova(LME_FRAP, type="III") ##site significant
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(Sqrt_test_FRAP~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)

# Cube 
# Make full model with treatment and site 
LME_FRAP <-lme(frap_cube1~Treatment+Site,random = ~1|Replicate, data=frap1)
Anova(LME_FRAP, type="III") ##site significant
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(frap_cube1~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)
```

# Run two-way ANOVAs to identify differences by site 
#### Comparing sites (MR, PR, UC, VP) and Treatments (MR-A, PR-UC-A, VP-A, Low) for Set2

```{r}
## FRAP - set 2

# First, I will do two-way ANOVA test ( using the untransformed variable)
# Make full model with treatment and site 
LME_FRAP <-lme(test_FRAP~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") 
#no significance for treatment or site, so do not move forward with any models 

# Second, I will do two-way ANOVA test using the transformed (sqrt and cube) variables to compare/check them to the untransformed
# Sqrt 
# Make full model with treatment and site 
LME_FRAP <-lme(Sqrt_test_FRAP~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") 
#no significance for treatment or site, so do not move forward with any models 

# Cube 
# Make full model with treatment and site 
LME_FRAP <-lme(frap_cube2~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") 
#no significance for treatment or site, so do not move forward with any models 
```

```{r}
## Delta FRAP - set 2

# First, I will do two-way ANOVA test ( using the untransformed variable)
# Make full model with treatment and site 
LME_FRAP <-lme(test_FRAP_Delta~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") # site significant 
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(test_FRAP_Delta~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)

# Second, I will do two-way ANOVA test using the transformed (log and sqrt) variables to compare/check them to the untransformed
# Log 
# Make full model with treatment and site 
LME_FRAP <-lme(Log_test_FRAP_Delta~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") # site significant 
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(Log_test_FRAP_Delta~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)

# Sqrt 
# Make full model with treatment and site 
LME_FRAP <-lme(Sqrt_test_FRAP_Delta~Treatment+Site,random = ~1|Replicate, data=frap2)
Anova(LME_FRAP, type="III") # site significant 
# Because only site was significant, make a reduced model with only site in it 
reduced_FRAP = lme(Sqrt_test_FRAP_Delta~Site,random = ~1|Replicate, data=frap1)
Anova(reduced_FRAP, type = "III") #site significant 
# Compare full and reduced model 
anova(LME_FRAP, reduced_FRAP)
##pvalue > 0.05, AIC differs by 4 = move forward with more parsimonious model aka reduced model 
summary(reduced_FRAP)
# Only site is significant so run Tukey post-hoc test on site only with reduced model
test <- glht(reduced_FRAP, linfct = mcp(Site = "Tukey"))
summary(test)
```

# Plot physiology variables variables 

```{r}
## FRAP - set 1
#Load data file with only avg and SE for each site. 
FRAP <- read.csv("data/FRAP/HT_FRAP_Siteavgs_043020.csv", header = T)
#gives me this warning when I load data file:
# Warning message:
#   In read.table(file = file, header = header, sep = sep, quote = quote,  :
#                   incomplete final line found by readTableHeader on '~/Desktop/Coding/R/HT/csv_files/FRAP/HT_FRAP_Siteavgs_043020.csv'
#not sure what it means, but did not affect plots

# Bar plot for FRAP set 1
f_p1 = ggplot(FRAP, aes(x=Site, y=FRAPavg1)) + #establishing what's going on the plot
  geom_bar(stat = "identity", color = "black", size = 0.3, width = 0.5) + # bar plot in these sizes, shapes, colors, etc
  geom_errorbar(aes(ymin = FRAPavg1 - FRAPse1, ymax = FRAPavg1 + FRAPse1), width = 0.2, position = position_dodge(0.75), size = 0.5) +
  ylab("Total antioxidant potential (μmol/g of tissue)") + xlab("") + # titles for x and y axes
  #ggtitle("FRAP set1" - Set1") + # figure title 
  scale_y_continuous(expand = c(0,0), limits = c(0, 1.35)) + #making the bars sit right on the x axis-- no white space bt the bars and the actual axis
  theme_minimal() + #setting the background theme
  theme(axis.line = element_line(), axis.text.x = element_text(size = 10, angle=30, hjust = 1, colour = "black"), axis.text.y = element_text(size=10, color = "black"), axis.title.y = element_text(size = 12)) #adding specific lines, sizes, degrees, etc for the axes texts and labels
f_p1
# no significance
```

```{r}
## FRAP - set 2
f_p2 <- ggplot(data = FRAP, aes(x=Site, y=FRAPavg2)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.75) +
  geom_errorbar(aes(ymin = FRAPavg2 - FRAPse2, ymax = FRAPavg2 + FRAPse2), width = 0.2, position = position_dodge(0.75), size = 0.5) +
  ylab("Total antioxidant potential (μmol/g of tissue)") + xlab("") + # titles for x and y axes
  #ggtitle("Total antioxidant potential (μmol/g of tissue) set2") +
  scale_y_continuous(expand = c(0,0)) + 
  theme_minimal() +
  theme(axis.line = element_line(), axis.text.x = element_text(size = 12, angle = 30, hjust = 1, colour = "black"), axis.text.y = element_text(size=12, colour = "black"), axis.title.y = element_text(size = 15))
f_p2
# no significance
```

```{r}
##delta FRAP
fDelta_p2 = ggplot(FRAP, aes(x=Site, y=dFRAPavg)) + # establishing whats going on in plot
  geom_bar(stat = "identity", color = "black", size = 0.3, width = 0.5) + # bar plot in these sizes, shapes, colors, etc
  geom_errorbar(aes(x=Site, ymin = dFRAPavg - dFRAPse, ymax = dFRAPavg + dFRAPse), colour = "black", width = 0.1, alpha = 0.9, size = 0.5) + #error bar in specific size, shape, color, etc
  geom_text(label = c("A", "B", "AB", "A"), aes(y = c(0.1, 0.05, 0.05, 0.21), x = Site), size = 4) + # adding letters to indicate significance 
  ylab("Change in total antioxidant potential (μmol/g of tissue)") + xlab("") + # x and y axes titles
  #ggtitle("Delta Condition Index") + # figure title 
  geom_hline(aes(yintercept = 0)) + # making the bars sit flush with the x axis
  scale_y_continuous(expand = c(0,0), limits = c(-0.4, 0.3)) + 
  theme_minimal() + # setting background theme
  theme(axis.line.y = element_line(), axis.text.x = element_text(size = 10, angle=30, hjust = 1, colour = "black"), axis.text.y = element_text(size=10, color = "black"), axis.title.y = element_text(size = 12)) #adding specific lines, sizes, degrees, etc for the axes texts and labels
fDelta_p2
# sig values: A, B, AB, A
```

```{r}
# Putting graphs into grid format by physiological variable
# FRAP - frap set1 and delta frap
frap_Plots <- ggarrange(f_p1, fDelta_p2, labels = c("A", "B"))
ggsave("output/FinalFigs/Figure5.png", plot = frap_Plots, width = 8, height = 4)
ggsave("output/FinalFigs/Figure5.pdf", plot = frap_Plots, width = 8, height = 4)
```







