---
title: "T-test"
author: "jillashey"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in data 
```{r}
phys <- read.csv("data/Glycogen_CI/HT_CIGly.csv", header = T)
head(phys)
phys$Site <- as.factor(phys$Site)
#set1 <- filter(phys, Set  == "Set1")
set2 <- filter(phys, Set  == "Set2")

stress <- read.csv("data/FRAP/HT_FRAP_bothSets_042320.csv", header = T)
head(stress)
stress$Site <- as.factor(stress$Site)
stress$Treatment <- as.factor(stress$Treatment)
#frap1 <- filter(stress, Set  == "Set1")
frap2 <- filter(stress, Set  == "Set2")
```

# Filter by site 
```{r}
# Phys df
MR <- filter(set2, Site  == "MR")
PR <- filter(set2, Site  == "PR")
UC <- filter(set2, Site  == "UC")
VP <- filter(set2, Site  == "VP")

# Stress phys df
MR.stress <- filter(frap2, Site  == "Merroir")
PR.stress <- filter(frap2, Site  == "Piankatank")
UC.stress <- filter(frap2, Site  == "Urbanna Creek")
VP.stress <- filter(frap2, Site  == "VIMS Pier")
```

Student's t-test for samples: use when you have one measurement variable and a theoretical expectation of what the mean should be under null. In this case, I am using t-tests to assess if the changes in phys. variables differed from 0
- Theoretical mean = 0; observed mean = avg of phys variable by site 
- Null hypothesis is that the mean change of phys variable is equal to 0
- If t-test is significant, can reject null--the mean change in phys. variable is significantly different from 0

### Only using t-tests for the change in phys. variables. We are comparing by site.

# T-test for physiology 
```{r}
##Set 2 - delta glycogen t-tests
#Do t-tests by site, as we will compare among sites 
t.test(MR$P_DeltaGly_DW) # not significant
t.test(PR$P_DeltaGly_DW) # significant 
t.test(UC$P_DeltaGly_DW) # not significant 
t.test(VP$P_DeltaGly_DW) # not significant 
# PR oysters were the only ones to significantly change from 0 during common garden conditions

##Set 2 - delta condition index t-tests
t.test(MR$Delta_CI) # significant
t.test(PR$Delta_CI) # significant
t.test(UC$Delta_CI) # significant
t.test(VP$Delta_CI) # not significant
#MR, PR, UC oysters all significantly changed from 0 during CG conditions 
```

# T-test for stress physiology 
```{r}
##Set 2 - delta FRAP t-tests
t.test(MR.stress$test_FRAP_Delta) # not significant
t.test(PR.stress$test_FRAP_Delta) # significant 
t.test(UC.stress$test_FRAP_Delta) # not significant 
t.test(VP.stress$test_FRAP_Delta) # not significant 
# PR oysters were the only ones to significantly change from 0 during common garden conditions
```





